---
output: github_document
params:
  iterations: 100
  input_bytes: 1000000
  output_csv: "benchmark-results/r-calls-rust.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
source("benchmarks/benchmark-core.R")
```

# RCallsRust

Small reproducible examples for comparing the overhead of R calling Rust through
several package styles. The examples deliberately do **not** depend on `sassy` or
any domain-specific crate. A pure C package is included as a baseline.

Each package exposes the same two functions:

- `count_byte(x, needle)`: count occurrences of one byte in a raw vector.
- `find_byte(x, needle)`: scan a raw vector and return a `data.frame` with
  zero-based positions and the matched byte.

The Rust kernel is intentionally simple so the examples focus on interface,
input access, and return-object construction overhead.

## Packages

- [`r/RCallsC`](r/RCallsC): pure C `.Call` implementation using R's native C API.
- [`r/RCallsRustC`](r/RCallsRustC): old-school `.Call` C wrapper plus a Rust
  `staticlib`, following the structure of
  [`r-rust/hellorust`](https://github.com/r-rust/hellorust).
- [`r/RCallsRustExtendrFfi`](r/RCallsRustExtendrFfi): an `extendr` package that
  uses `extendr_ffi` directly for raw-vector access and data-frame construction.
  R wrappers are generated from `extendr` metadata.
- [`r/RCallsRustExtendr`](r/RCallsRustExtendr): a high-level `extendr` package
  using `Raw` and `data_frame!`. R wrappers are generated from `extendr`
  metadata.
- [`r/RCallsRustSavvy`](r/RCallsRustSavvy): a
  [`savvy`](https://github.com/yutannihilation/savvy) package variant. Its C/R
  wrappers are generated with `savvy::savvy_update()`. It returns a named list
  from Rust and constructs the final `data.frame` in R, which matches savvy's
  documented recommendation for data frames.

All `find_byte()` implementations use the same two-pass algorithm: first count
matches to determine output size, then scan again to fill positions. This keeps
the Rust-backed variants algorithmically aligned with the C/R API variants,
which need the output size before allocating R vectors.

## Make targets

From the repository root:

```sh
make deps      # install tinytest, bench, ggplot2, and rmarkdown
make install   # install all packages
make test      # run tinytest for all packages
make bench     # run bench::mark benchmark script
make report    # render README.md and benchmarks/benchmark.md
# Linux benchmarks pin to one CPU by default; use RCALLSRUST_BENCH_CPU=none to disable.
# Use RCALLSRUST_BENCH_CPU=<cpu> to choose a specific CPU.
make check     # run R CMD check for all packages
```

## Benchmark results

The benchmark uses the R `bench` package, a deterministic 1 MB raw vector by
default, and reports both scalar count and small-data-frame return timings. The
quantiles are included because these sub-millisecond scans can have visible
scheduler/cache tails. Each case is measured in a separate `bench::mark()` run
to reduce cross-case cache and allocation interference. Timing columns are
reported in microseconds; `itr_per_second` is bench's aggregate throughput and is not simply
`1e6 / median_us`. The distribution plot uses a log10 microsecond scale so
long-tail outliers do not flatten the main timing bands. On Linux, benchmark
runs pin themselves to one CPU by default (`RCALLSRUST_BENCH_CPU=auto`) to avoid
core-migration tails on heterogeneous or frequency-scaled CPUs. Set
`RCALLSRUST_BENCH_CPU=<cpu>` to choose a CPU, or `RCALLSRUST_BENCH_CPU=none` to
disable pinning.

```{r benchmark, echo=FALSE}
bench_iterations <- param_int("RCALLSRUST_BENCH_ITERS", params$iterations)
bench_input_bytes <- param_int("RCALLSRUST_BENCH_N", params$input_bytes)
bench <- get_rcallsrust_benchmark(
  n = bench_input_bytes,
  iterations = bench_iterations,
  output_csv = params$output_csv
)
knitr::kable(bench$input)
knitr::kable(bench$results, digits = 6)
```

```{r benchmark-plot, echo=FALSE, fig.width=11, fig.height=6}
plot_rcallsrust_benchmark(bench$mark, type = "boxplot", log_scale = TRUE)
```

The CSV artifact is written to `r bench$output_csv`; the raw `bench::mark()`
object used for the plot is written to `r bench$output_rds`.

A standalone benchmark report is also available at
[`benchmarks/benchmark.md`](benchmarks/benchmark.md), generated from
[`benchmarks/benchmark.Rmd`](benchmarks/benchmark.Rmd).
