---
title: "RCallsRust benchmark"
output:
  github_document:
    toc: true
    toc_depth: 2
params:
  iterations: 100
  input_bytes: 1000000
  output_csv: "benchmark-results/r-calls-rust.csv"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
source("benchmark-core.R")
```

This benchmark compares five small R packages that all scan the same raw vector
and expose the same R API:

- `RCallsC`: pure C `.Call` implementation.
- `RCallsRustC`: C `.Call` wrapper plus a Rust static library.
- `RCallsRustExtendrFfi`: `extendr` registration with direct `extendr_ffi` use.
- `RCallsRustExtendr`: high-level `extendr` wrapper and `data_frame!` output.
- `RCallsRustSavvy`: `savvy` wrappers, returning a named list from Rust and
  converting to a data frame in R.

The benchmark uses the R `bench` package. All `find_byte()` implementations use
the same two-pass algorithm: count matches, allocate output, then fill positions.
Quantiles are reported because these sub-millisecond scans can have visible
scheduler/cache tails; `itr_per_second` is bench's aggregate throughput and is
not simply `1 / median_seconds`.

## Machine

```{r machine, echo=FALSE}
machine <- data.frame(
  field = c("system", "release", "machine", "R", "platform", "cargo", "rustc"),
  value = c(
    unname(Sys.info()[["sysname"]]),
    unname(Sys.info()[["release"]]),
    unname(Sys.info()[["machine"]]),
    R.version.string,
    R.version$platform,
    paste(system2("cargo", "--version", stdout = TRUE), collapse = "\n"),
    paste(system2("rustc", "--version", stdout = TRUE), collapse = "\n")
  )
)
knitr::kable(machine)
```

## Results

```{r benchmark, echo=FALSE}
repo_root <- normalizePath("..", mustWork = TRUE)
output_csv <- params$output_csv
if (!grepl("^/|^[A-Za-z]:[/\\\\]", output_csv)) output_csv <- file.path(repo_root, output_csv)

bench_iterations <- param_int("RCALLSRUST_BENCH_ITERS", params$iterations)
bench_input_bytes <- param_int("RCALLSRUST_BENCH_N", params$input_bytes)
bench <- get_rcallsrust_benchmark(
  n = bench_input_bytes,
  iterations = bench_iterations,
  output_csv = output_csv
)
knitr::kable(bench$input)
knitr::kable(bench$results, digits = 6)
```

## Plot

```{r benchmark-plot, echo=FALSE, fig.width=9, fig.height=6}
plot_rcallsrust_benchmark(bench$mark, type = "boxplot")
```

The CSV artifact is written to `r bench$output_csv`; the raw `bench::mark()`
object used for the plot is written to `r bench$output_rds`.
